\bf Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-07-15 manuscript/main.tex
right before the proposition 3

One possible approach is to extend smoothed functional PCA \citep{Rice1991} to our optimization problem \eqref{def:maximizer_squared_empirical_cov}, leading to the following formulation:
	\begin{equation*}
	\hat{\xi} 
	= 
	(\hat{\xi}_1, \ldots, \hat{\xi}_K, \hat{\boldsymbol{\xi}})
	= \arg \max_{\xi \in \mathcal{B}^K \times \mathbb{R}^p,
		\| \xi  \|_\mathcal{H} = 1
	} 
	\biggl\{
	\widehat{\textnormal{Cov}}^2 
	\bigl(
	\langle \widetilde{W}, \xi \rangle_{\mathcal{H}}, Y 
	\bigr)
	- \sum \limits_{j=1}^K \lambda_j 
	\int_0^1 \bigl\{ \hat{\xi}_j^{\prime\prime}(t) \bigr\}^2 dt
	\biggr\}
	,
\end{equation*}
where the smoothing parameters $\{\lambda_j\}_{j=1}^K$ control the trade-off between covariance maximization and roughness penalization.
However,   we take an alternative strategy that extends the idea of \citet{Silverman1996}. This approach is known to produce appropriately smoothed functional PCA components under milder conditions than the \citet{Rice1991}'s approach by replacing the usual $L^2$-orthonormality constraint with the one that takes account the roughness of functions via the modified inner product. Specifically, the following proposition provides a generalized Rayleigh quotient problem which modifies the $\mathcal{H}$-orthonormality constraint of the hybrid PLS components to incorporate roughness penalty, and thus whose solution corresponds to a smoothed PLS component maximizing $\widehat{\textnormal{Cov}}^2 \left(\langle \mathbf{W}, \xi \rangle_\mathcal{H}, Y \right)$ over the class of all functions satisfying sufficient smoothness.
	\begin{proposition} \label{Proposition: Regularized generalized rayleigh quotient}
		At each $l$-th iteration of the first step of the PLS algorithm involving the residualized response and predictors $Y_i \equiv Y_i^{[l]}$ and $\mathbf{W}_i \equiv \mathbf{W}_i^{[l]}$, consider the basis function expansions of the functional predictors $X_i^{(k)}(t_k) = \sum_{m=1}^M c_{im}^{(k)} b^{(k)}_{m}(t_k)$, $i=1, \ldots, n$, and those of the functional parts of the PLS component $\psi^{(k)}(t_k) = \sum_{m=1}^M a^{(k)}_{m} b^{(k)}_m(t_k)$, $k=1, \ldots, K$. Recall from \eqref{def:V_star} that  $V^\ast$ is defined as
		\begin{equation*}
			\mathbf{V}^\ast = n^{-2}\begin{bmatrix}
				J \widetilde{C}^\top \mathbf{y}\mathbf{y}^\top \widetilde{C} J & J \widetilde{C}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \mathbf{Z} \\
				\mathbf{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C} J &   \mathbf{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \mathbf{Z}
			\end{bmatrix} = n^{-2} J^\ast \widetilde{C}^{*\top} \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C}^{*} J^\ast
			\in \mathbb{R}^{(MK+p) \times (MK+p)}.
		\end{equation*}
		and  $\xi^*=(\mathbf{a}^{\top}, \boldsymbol{\theta}^{\top})^\top \in \mathbb{R}^{MK + p}$. Then, the $l$-th smoothed PLS component, $\widehat{\xi} \equiv \widehat{\xi}_l$ can be obtained in the following way.
		We first compute the basis coefficients by solving the optimization problem
		Note that the penalty term can be written concisely using matrices as:
		\begin{equation}
			\textnormal{PEN} \left( \hat{\xi}_j \right) 
			=  \int_0^1 
			\bigl\{ \hat{\xi}_j^{''}(t) \bigr\}^2 dt = \mathbf{a}^{(k)T} \ddot{J}^{(k)} \mathbf{a}^{(k)},
			\label{eq: roughness penalty}
		\end{equation}
		where the last term is obtained using the basis expansion $\psi^{(k)}(t)=\mathbf{a}^{(k)T}\mathbf{b}^{(k)}(t)$.

>>>>

One possible approach is to extend smoothed functional PCA \citep{Rice1991} to our optimization problem \eqref{def:maximizer_squared_empirical_cov}, leading to the following formulation:
	\begin{equation*}
	\hat{\xi} 
	= 
	(\hat{\xi}_1, \ldots, \hat{\xi}_K, \hat{\boldsymbol{\xi}})
	= \arg \max_{\xi \in \mathcal{B}^K \times \mathbb{R}^p,
		\| \xi  \|_\mathcal{H} = 1
	} 
	\biggl\{
	\widehat{\textnormal{Cov}}^2 
	\bigl(
	\langle \widetilde{W}, \xi \rangle_{\mathcal{H}}, Y 
	\bigr)
	- \sum \limits_{j=1}^K \lambda_j 
	\int_0^1 \bigl\{ \hat{\xi}_j^{\prime\prime}(t) \bigr\}^2 dt
	\biggr\}
	,
\end{equation*}
where the smoothing parameters $\{\lambda_j\}_{j=1}^K$ control the trade-off between maximizing covariance and penalizing roughness. However, the above approach assumes that the functional predictors admit an orthogonal expansion in the $L^2$ sense. As an alternative, we adopt the strategy of \citet{Silverman1996}, which works under weaker conditions than those required by \citet{Rice1991}. This is achieved by replacing the standard orthonormality constraint with one based on a modified inner product that accounts for roughness. Building on this idea, the following proposition reformulates the hybrid PLS problem as a generalized Rayleigh quotient, yielding a smoothed PLS direction that maximizes the squared covariance over a class of smooth functions.

	\begin{proposition} \label{Proposition: Regularized generalized rayleigh quotient}
		At each $l$-th iteration of the first step of the PLS algorithm involving the residualized response and predictors $Y_i \equiv Y_i^{[l]}$ and $\mathbf{W}_i \equiv \mathbf{W}_i^{[l]}$, consider the basis function expansions of the functional predictors $X_i^{(k)}(t_k) = \sum_{m=1}^M c_{im}^{(k)} b^{(k)}_{m}(t_k)$, $i=1, \ldots, n$, and those of the functional parts of the PLS component $\psi^{(k)}(t_k) = \sum_{m=1}^M a^{(k)}_{m} b^{(k)}_m(t_k)$, $k=1, \ldots, K$. Recall from \eqref{def:V_star} that  $V^\ast$ is defined as
		\begin{equation*}
			\mathbf{V}^\ast = n^{-2}\begin{bmatrix}
				J \widetilde{C}^\top \mathbf{y}\mathbf{y}^\top \widetilde{C} J & J \widetilde{C}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \mathbf{Z} \\
				\mathbf{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C} J &   \mathbf{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \mathbf{Z}
			\end{bmatrix} = n^{-2} J^\ast \widetilde{C}^{*\top} \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C}^{*} J^\ast
			\in \mathbb{R}^{(MK+p) \times (MK+p)}.
		\end{equation*}
		and  $\xi^*=(\mathbf{a}^{\top}, \boldsymbol{\theta}^{\top})^\top \in \mathbb{R}^{MK + p}$. Then, the $l$-th smoothed PLS component, $\widehat{\xi} \equiv \widehat{\xi}_l$ can be obtained in the following way.
		We first compute the basis coefficients by solving the optimization problem
		Note that the penalty term can be written concisely using matrices as:
		\begin{equation}
			\textnormal{PEN} \left( \hat{\xi}_j \right) 
			=  \int_0^1 
			\bigl\{ \hat{\xi}_j^{''}(t) \bigr\}^2 dt = \mathbf{a}^{(k)T} \ddot{J}^{(k)} \mathbf{a}^{(k)},
			\label{eq: roughness penalty}
		\end{equation}
		where the last term is obtained using the basis expansion $\psi^{(k)}(t)=\mathbf{a}^{(k)T}\mathbf{b}^{(k)}(t)$.
